Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2024-01-13 src/tex/ms.tex
added response 4

Aligned-spin, frequency-domain waveform models (such as IMRPhenomD) can be written as a
combination of amplitude and phase functions ($A$ and $\phi$ respectively):
\begin{align}\label{eq:}
	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda^i)}\,,
\end{align}
where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda^i$ \corr{($i=1,2,\cdots,19$) is a set of 19 additional
parameters} which will be discussed below. 
The phase and amplitude functions are then split into three sections which represent the
inspiral, intermediate, and merger-ringdown (MR) parts of the waveform. 
During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up 
to 3.5PN order \corr{augmented with the next four higher order PN terms}.
To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
uses a series of parameterizations\footnote{
	The parameterizations for both the amplitude and phase functions can be found in \citep{khan2016frequency}.
} 
which depend purely on $\Lambda^i$ and can be calibrated to numerical relativity (NR) simulations.
The three sections are then \textit{stitched} together using step functions.
Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
boundary between each section.

@@

The correlation we observe between the mismatches and the waveform's parameters
can also be used to design more physical ansatz. For example, we see the
recalibration process struggles to improve in the regime where $\chi_{1}$ is
positive and $\chi_{2}$ is negative. \corr{\sout{This means there is a dependency between
the two spins} \sout{that is not captured by the current ansatz.}} A systematic scheme that
incorporates our calibration method with waveform design will be handy in
constructing new waveform models in the future.

\corr{We note that the parameter space we use is restricted to $q\leq8$, while waveforms 
of $q=18$ are used in the original calibration trials. Also, we have directly used NR 
waveforms instead of hybridized SEOB-NR waveforms for training and validation purposes. 
These choices put the original IMRPhenomD result at a disadvantage, as both the parameter 
space and frequency range are restricted. We tested our result by training the model 
using a set of training waveforms of the same intrinsic parameters, but with BAM waveforms 
replaced with SEOBNR waveforms. We can see the result in Fig.~\ref{fig:q14818}, where the 
mismatch median decreases by about $8\%$, This mismatch improvement mainly concerns the 
stitching problem discussed before. Nevertheless, it is known that IMRPhenomD has a large 
inaccuracy in the large mass-ratio regime, and is rarely used in this region of parameter 
space. Hence, we focus on showing AD with gradient descent, as a general and robust method, 
can be used to further optimize GW waveform models with arbitrary sets of training 
waveforms used.}

@@

In making Fig.~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions 
purely for simplicity. These simple cuts demonstrate the accuracy of the waveform can be 
further improved, but they are almost certainly not the optimal way to incorporate the 
extra information we have about the waveform. A more general way such as adding new 
functional dependency between the parameters should be explored in the future.

>>>>

Aligned-spin, frequency-domain waveform models (such as IMRPhenomD) can be written as a
combination of amplitude and phase functions ($A$ and $\phi$ respectively):
\begin{align}\label{eq:}
	h(f,\theta,\Lambda^i) = A(f,\theta,\Lambda^i)e^{-i\phi(f,\theta,\Lambda^i)}\,,
\end{align}
where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda^i$ \corr{($i=1,2,\cdots,19$) is a set of 19 additional
parameters} which will be discussed below. 
The phase and amplitude functions are then split into three sections which represent the
inspiral, intermediate, and merger-ringdown (MR) parts of the waveform. 
During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up 
to 3.5PN order \corr{augmented with the next four higher order PN terms}.
To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
uses a series of parameterizations\footnote{
	The parameterizations for both the amplitude and phase functions can be found in \citep{khan2016frequency}.
} 
which depend purely on $\Lambda^i$ and can be calibrated to numerical relativity (NR) simulations.
The three sections are then \textit{stitched} together using step functions.
Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
boundary between each section.

@@

The correlation we observe between the mismatches and the waveform's parameters
can also be used to design more physical ansatz. For example, we see the 
recalibration process struggles to improve in the regime where $\chi_{1}$ is 
positive and $\chi_{2}$ is negative. This means there is a dependency between the two 
spins that is not captured in the current ansatz. \corr{In fact, this issue can 
be explained by PN theory, where the leading spin-orbit coupling term depends on 
the effective spin while higher order terms depends on both $\chi_1$ and $\chi_2$. 
Nevertheless, we hope to show a systematic scheme that could reveal hidden degeneracies, 
as illustrated by the effective spin degeneracy. Hence, such systematic scheme that 
incorporates our calibration method with waveform design will be handy in constructing 
new waveform models in the future. }

\corr{We note that the parameter space we use is restricted to $q\leq8$, while waveforms 
of $q=18$ are used in the original calibration trial. Also, we have directly used NR 
waveforms instead of hybridized SEOB-NR waveforms for training and validation purposes. 
These choices put the original IMRPhenomD result at a disadvantage, as both the parameter 
space and frequency range are restricted. We tested our result by training the model 
using a set of training waveforms of the same intrinsic parameters, but with BAM waveforms 
replaced with SEOBNR waveforms. We can see the result in Fig.~\ref{fig:q14818}, where the 
mismatch median decreases by about $8\%$, This mismatch improvement mainly concerns the 
stitching problem discussed before. Nevertheless, it is known that IMRPhenomD has a large 
inaccuracy in the large mass-ratio regime, and is rarely used in this region of parameter 
space. Hence, we focus on showing AD with gradient descent, as a general and robust method, 
can be used to further optimize GW waveform models with arbitrary sets of training 
waveforms used.}

@@

In making Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions 
purely for simplicity. These simple cuts demonstrate the accuracy of the waveform can be 
further improved, but they are almost certainly not the optimal way to incorporate the 
extra information we have about the waveform. A more general way such as adding new 
functional dependency between the parameters should be explored in the future.
